 \stepcounter{lecture}
 \setcounter{lecture}{8}
 \sektion{Lecture 8}

\begin{proof}[Valuative Criterion proof continued]
still in the works.
\end{proof}

\begin{corollary}
When working with noetherian schemes: \marginpar{Separatedness on
noetherian schemes}
\begin{itemize}
 \item[(a)] Open and closed immersions are separated.
 \item[(b)] Compositions of separated morphisms are separated.
 \item[(c)] Separatedness is stable under base extension. i.e.  if
 $f:X\to Y$ is separated, then $f':X\times_Y Y' \to Y'$ is
 separated.
 \[\xymatrix{
 X\times_Y Y' \ar[r] \ar[d]^{f'} & X\ar[d]^f\\
 Y' \ar[r] & Y
}\]
 \item[(d)] If $f:X\to Y$ and $f':X'\to Y'$ are separated
 $S$-morphisms, then so is $f\times f':X\times_S X' \to Y\times_S
 Y$.
 \item[(e)] If $X\xrightarrow{f} Y \xrightarrow{g} Z$ are
 morphisms and $g\circ f$ is separated, then so is $f$.
 \item[(f)] Separatedness is local on the base.  i.e. $f:X\to Y$
 is separated if and only if there is an open cover $\{U_i\}$ of
 $Y$ such that $f^{-1}(U_i)\to U_i$ is separated for all $i$.
 \item[(g)] $f:X\to Y$ is separated if and only if
 $f_{\text{red}}:X_{\text{red}}\to Y_{\text{red}}$ is separated.
\end{itemize}
\end{corollary}
\begin{proof}[Proof of (e)]
in the works.
\end{proof}
\begin{proof}[Proof of (g)]
in the works.
\end{proof}

\begin{remark}
All of these things are true without the noetherian hypothesis.
\end{remark}

\begin{corollary}[of the Corollary]
A morphism of separated schemes is separated.  More precisely, if
$f:X\to Y$ with $X$ separated (over $\Z$), then $f$ is separated.
\end{corollary}
\begin{proof}
Use (e) of the Corollary above.
\end{proof}

\begin{corollary}[of the Corollary]
Most schemes you work with will be separated.  An exception:
\underline{gluing}.
\end{corollary}

\marginpar{Properness} In topology, a map is proper if the inverse
image of a compact set is compact.  We want $\P^1_k\to \spec k$ to
be proper, but $\mathbb{A}^1_k\to \spec k$ not to be proper.

\begin{definition}
A morphism $f:X\to Y$ is \emph{closed} if for all closed
subschemes $Z\subseteq X$, $f(Z)$ is closed in $Y$.  We say that
$f$ is \emph{universally closed} if for all morphisms $Y'\to Y$,
the pullback of $f$ is closed.
\[\xymatrix{
 X'\ar[d]^{f'} \ar[r] & X\ar[d]^f\\
 Y'\ar[r] & Y
}\]
\end{definition}

Example: $\mathbb{A}^1_k\to \spec k$ is closed, but not
universally closed.  Take $Y'= \mathbb{A}^1_k$, so $X' =
\mathbb{A}^2_k$, and $f'$ is a projection.  Then look at
$Z=\{xy=1\}$.

\begin{definition}
A morphism is \emph{proper} if it is separated, of finite type,
and universally closed.
\end{definition}

\begin{theorem}[Valuative Criterion for Properness]
\marginpar{Valuative Criterion for Properness} Let $f:X\to Y$ be a
finite type morphism of noetherian schemes, then $f$ is proper if
and only if
\[\xymatrix{
 U:= \spec K \ar[r] \ar[d]^i & X\ar[d]^f\\
 T:= \spec R \ar[r] \ar@{.>}[ur]^{\exists !} & Y
}\]
\end{theorem}
\begin{proof}
in the works.
\end{proof}
